The Galactic Delay-Time Distribution of Type Ia Supernovae:\\ A Chemical Evolution Perspective

### 2022-08-10 src/tex/ms.tex
[showyourwork] overleaf sync

% showyourwork!
\usepackage{showyourwork}

% Units in equations
% \usepackage{siunitx}

@@

% Title
\title{An open source scientific article}

% Author list
\author[0000-0003-3781-0747]{Liam O. Dubay}

@@

\begin{figure}
    \centering
    \includegraphics{figures/onezone_slope_timescale.pdf}
    \caption{\textit{Center:} Abundance tracks in the [Fe/H]-[O/Fe] plane for one-zone models with various DTD parameters. Tracks from models which assume a power-law DTD are in black, with dotted, dashed, and solid curves representing power-law slopes ($\alpha$) of $-0.8$, $-1.1$, and $-1.4$, respectively. Tracks from models which assume an exponential DTD are in purple, with dotted, dashed, and solid curves representing exponential decay timescales ($\tau$) of 1.5 Gyr, 3 Gyr, and 6 Gyr, respectively. In all cases a minimum delay time of 40 Myr is assumed. \textit{Top:} MDFs generated from by one-zone models. \textit{Right:} Distribution of [O/Fe] generated by the one-zone models.}
    \label{fig:onezone-slope-timescale}
    \script{onezone_slope_timescale.py}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{figures/onezone_dtd_delay.pdf}
    \caption{Similar to Figure \ref{fig:onezone-slope-timescale} but with varying minimum SN Ia delay time ($t_D$). Dashed curves represent models with a minimum delay time of 40 Myr, while solid curves represent models with a minimum delay time of 150 Myr. The black curves represent a power-law DTD with a slope of $-1.1$, while the purple curves represent an exponential DTD with a timescale of 1.5 Gyr.}
    \label{fig:onezone-dtd-delay}
    \script{onezone_dtd_delay.py}
\end{figure}

@@

>>>>

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{siunitx}

% custom unit definitions
\DeclareSIUnit\year{yr}

@@

% Title
\title{The Galactic Delay-Time Distribution of Type Ia Supernovae:\\
    A Chemical Evolution Perspective}

% Author list
\author[0000-0003-3781-0747]{Liam O. Dubay}
\author[0000-0001-7258-1834]{Jennifer A. Johnson}
\author[0000-0002-6534-8783]{James W. Johnson}
% \author[0000-0001-7775-7261]{David H. Weinberg}

@@

\begin{figure}
    \centering
    \includegraphics{figures/onezone_slope_timescale.pdf}
    \caption{\textit{Center:} Abundance tracks in the [Fe/H]-[O/Fe] plane for one-zone models with various DTD parameters. Tracks from models which assume a power-law DTD are in black, with dotted, dashed, and solid curves representing power-law slopes ($\alpha$) of $-0.8$, $-1.1$, and $-1.4$, respectively. Tracks from models which assume an exponential DTD are in blue, with dotted, dashed, and solid curves representing exponential decay timescales ($\tau$) of 1.5 Gyr, 3 Gyr, and 6 Gyr, respectively. In all cases a minimum delay time of 40 Myr is assumed. \textit{Top:} MDFs generated from by one-zone models. \textit{Right:} Distribution of [O/Fe] generated by the one-zone models.}
    \label{fig:onezone-slope-timescale}
    \script{onezone_slope_timescale.py}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{figures/onezone_dtd_delay.pdf}
    \caption{Similar to Figure \ref{fig:onezone-slope-timescale} but with varying minimum SN Ia delay time ($t_D$). Dashed curves represent models with a minimum delay time of 40 Myr, while solid curves represent models with a minimum delay time of 150 Myr. The black curves represent a power-law DTD with a slope of $-1.1$, while the blue curves represent an exponential DTD with a timescale of 1.5 Gyr.}
    \label{fig:onezone-dtd-delay}
    \script{onezone_dtd_delay.py}
\end{figure}

@@

\subsubsection{Analytical Formulations}

\citet{Greggio2005-AnalyticalRates} derives analytical DTDs for SD and DD progenitor systems from assumptions about binary stellar evolution and outcomes of mass exchange. They find that the parameters which have a large effect on the shape of the DTD are the distribution and range of stellar masses in progenitor systems; the efficiency of accretion in the SD scenario; and the distribution of separations at birth in the DD scenario. Figure \ref{fig:greggio05-dtd} shows the analytical DTDs for SD progenitors and two different prescriptions for DD progenitors (``WIDE'' and ``CLOSE''). In the ``WIDE'' scheme, it is assumed that there is a wide distribution of ratios $A/A_0$ of the separation of the DD system to the initial separation of the binary, and that the distributions of $A$ and total mass of the system $m_{\rm{DD}}$ are independent, so one cannot necessarily predict the total merge time of a system based on its initial parameters. In the ``CLOSE'' scheme, there is assumed to be a narrow distribution of $A/A_0$ and a correlation between $A$ and $m_{\rm{DD}}$, so the most massive binaries tend to merge quickly and the least massive merge last.

An implementation of these DTDs in a chemical evolution model requires numerical integration (see their Equations 14, 33, and 34, respectively, for the exact functional forms), so to save on computation time we approximate the DTD for the DD prescriptions as a broken power-law with an exponential rise and tail:

\begin{equation}
    f_{\rm{Ia}}^{\rm{DD}}(t) \propto
    \begin{cases}
        t^{\alpha_1} (1 - A_{\rm{rise}} e^{-t/\tau_{\rm{rise}}}) & t \le \SI{1}{\giga\year} \\
        B t^{\alpha_2} (1 + A_{\rm{tail}} e^{-(t - \SI{1}{\giga\year})/\tau_{\rm{tail}}}) & t > \SI{1}{\giga\year}
    \end{cases}
    \label{eq:greggio05-approx}
\end{equation}

\noindent with the normalization $B$ of the second part set such that

\begin{equation}
    B = \frac{1}{1 + A_{\rm{tail}}} t^{(\alpha_1-\alpha_2)} (1 - A_{\rm{rise}} e^{-(\SI{1}{\giga\year})/\tau_{\rm{rise}}}).
\end{equation}

\noindent We use a non-linear least squares method to fit Equation \ref{eq:greggio05-approx} to the analytical DTDs for DD progenitors. We obtain best-fit parameters $\alpha_1=-0.19$, $A_{\rm{rise}}=1.7$, $\tau_{\rm{rise}}=0.084$ Gyr, $\alpha_2=-0.91$, $A_{\rm{tail}}=0.85$, and $\tau_{\rm{tail}}=0.088$ Gyr for the ``WIDE'' scheme, and $\alpha_1=-0.53$, $A_{\rm{rise}}=3.0$, $\tau_{\rm{rise}}=0.042$ Gyr, $\alpha_2=-1.1$, $A_{\rm{tail}}=0.62$, and $\tau_{\rm{tail}}=0.11$ Gyr for the ``CLOSE'' scheme. These best-fit approximations are represented by the dashed curves in Figure \ref{fig:greggio05-dtd}. The difference between the true analytical DTDs and our best-fit approximations is small enough as to be irrelevant for our purposes, so for the rest of this section we rely on the best-fit approximations for faster simulations.

\begin{figure}
    \centering
    \includegraphics{figures/greggio05_dtd.pdf}
    \caption{Analytical DTDs computed by \citet{Greggio2005-AnalyticalRates}.}
    \label{fig:greggio05-dtd}
    \script{greggio05_dtd.py}
\end{figure}

Figure \ref{fig:onezone-greggio05} shows the results of one-zone chemical evolution models with the \citet{Greggio2005-AnalyticalRates} DTDs. We assume $\eta=2.5$, $\tau_*=2$ Gyr, an inside-out SFH evaluated at a radius of 8 kpc, a continuous recycling approximation, and a minimum SN Ia delay of 40 Myr. We compare the SD, DD WIDE, and DD CLOSE schemes to standard power-law and exponential DTDs. The SD and DD CLOSE DTDs follow nearly identical tracks in [O/Fe] vs [Fe/H] to an exponential DTD with a 1.5 Gyr timescale. The DD WIDE DTD likewise follows a similar track to a 3 Gyr exponential.

\begin{figure}
    \centering
    \includegraphics{figures/onezone_greggio05.pdf}
    \caption{Caption}
    \label{fig:onezone-greggio05}
    \script{onezone_greggio05.py}
\end{figure}
